### 2020-11-19 paper/ifo-tableTopGWs.tex

\documentclass[paper-main.tex]{subfiles}

\begin{document}

% rephrased to mention the laser as the first component and shifted the arm lengths of LIGO and Virgo to later in the para when we discuss them. 
Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. 
However their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern, 
The resulting interference pattern is dependent on the relative distance travelled by the beams. 
Current generation gravitational wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
They are able to measure minuscule changes in distance due to gravitational waves, e.g. the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton. 
%changed the relative length of the two arms by $10^{-21}$, a distance smaller than the width of a proton~\cite{GW150914}.

\begin{comment}
Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. However, their design is based fundamentally on a Michelson interferometer, an optical configuration commonly used in undergraduate laboratories.
It consists of a beam splitter which splits light from a laser down two perpendicular arms. % I suggest moving this to later as we are talking about Michelson interferometers generally here. 
%($4\,{\rm km}$ and $3\,{\rm km}$ long at LIGO and Virgo respectively).
Mirrors at the end of each arm reflect the light back towards the beam splitter, where they recombine and produce an interference pattern.
This pattern is dependent on the relative distance travelled by each beam and thus on the relative length of each arm. 
Current generation gravitational wave observatories are able to measure minuscule changes in distance due to passing gravitational waves. 
The first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
\end{comment}

% I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
Sound is a commonly-used analogy when explaining gravitational wave science. 
Gravitational wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot. Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
% rephrased as this what we are doing, rather than what scientists in general are doing
%This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer. 
Detection and analysis techniques used by the gravitational wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
We use audio signals throughout this work to simulate gravitational wave-like signals.
% and  to demonstrate different data analysis and signal processing techniques used to recover such signals.

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam_edit.pdf}
	\caption{\label{fig:ifo_schematic_webcam}
Schematic of the table-top Michelson interferometer. 
The output of a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO} (top-left) is reflected from a mirror at a right angle. The beam then passes through a converging lens (FL = $125.0\,{\rm mm}$) before reaching the beamsplitter which is $50$/$50$ in reflection and transmission. 
The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 
The output interference pattern is enlarged using a diverging lens (FL = $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a computer (PC). 
A speaker fixed to the back of M2 is used to inject audio signals from the PC into the interferometer.
    }
\end{figure}

The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green eye-safe laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
After passing through a converging (plano-convex) lens with focal length $125.0\,{\rm mm}$, the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2). 
A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a PC. 
The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
The beams are reflected by mirrors M1 and M2, located at the end of $\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long arms.
The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of focal length $-25.0\,{\rm mm}$, and projected onto a screen.
The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
A change in the relative length of the arms causes these rings to move radially in or outwards.
Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
% Ring interference patterns like this are often used in gravitational wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
% Why remove the sentence above? Rings are chosen rather than happening by default. - Hannah
The intensity timeseries is recorded via webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
In Section~\ref{sec:optical_microphone}, which considers more complex audio, the webcam is replaced with a photodiode, which offers a higher sample rate than the webcam. 
%rephrased as this sentence is confusing I think: An intensity time series of some point in the interference pattern is recorded by a webcam (or a photodiode in Section~\ref{sec:optical_microphone}) for data analysis to extract the signal injected at the speaker.

\begin{figure}
 \begin{center}
  \includegraphics[height=0.3\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
 \end{center}
 \caption{\label{fig:interference_pattern}
The interference pattern produced by the table-top Michelson interferometer.
This image was taken with the webcam off the beam axis, which may explain why the (predicted circular) pattern appears elliptical.
The pink star indicates the point where the intensity time series measurements were taken.
The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
}
\end{figure}

The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
The amplitude of these motions is given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
If the relative change in the length of the arms is kept small enough, then the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
But if the relative change in the length of the arms is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.
As such, any motion of the fringes must be kept small by playing the sound softly through the speaker.
% too much to have --- and () in the same sentence I think 
Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena like frequency doubling.
We derive this nonlinear relationship between the length difference of the arms and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.

% removed as almost identical to the next sentence
%In the following section, we start simply by injecting and recovering a audio signal of constant frequency (a single tone).

% Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photodiode. 
% In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
% In Section~\ref{sec:optical_microphone}, which considers more complex audio, we turn to a photodiode because it offers a higher sample rate than the webcam.

\end{document}